---
title: "Machine learning for policy analysis"
#author: "Dr. Stephan Dietrich & Michelle González Amador"
#date: '2022-09-21'
output: html_document
---

At its core, machine learning is primarily interested in making sense of complex data. A machine learning algorithm's job is to read data and identify patterns that can be used for: prediction, feature-relevance detection, model-free classification, among other actions. 

## **A general introduction** {.tabset}

### **Machine Learning for Public Policy**:


Empirical Public Policy research is primarily concerned with causality and understanding counterfactuals. What is the impact of Policy or Programme X on Population Outcome Y? What would have happened in the absence of such a policy? However, there are policy problems that may be solved without necessarily looking for a causal link. We call these `prediction policy problems' ([Kleinberg et al. 2015](https://www.aeaweb.org/articles?id=10.1257/aer.p20151023)). Some examples include: 


+ Allocating fire and health inspectors in cities ([Glaeser et al.    2016](https://www.aeaweb.org/articles?id=10.1257/aer.p20161027)).


+ Predicting unemployment spell length to help workers decide on savings rates and job search strategies.


+ Predicting highest risk youth for targeting interventions ([Chandler,Levitt, and List 2011](https://www.aeaweb.org/articles?id=10.1257/aer.101.3.288)).


**How does a Machine Learn?**


Machine learning is classified in three major branches: 

:::: {style="display: grid; grid-template-columns: 1fr 1fr 1fr; grid-column-gap: 8px"}

::: {}


**Supervised Learning**:


This course will be primarily concerned with supervised learning. Supervised learning is analogous to statistical learning: suppose that we observe a quantitative response $Y$ and $p$ predictors, $X_1, X_2,..., X_p$.


We can rewrite this in a general linear form as:


$$Y = f(X) + u$$
where $u$ is an error term. 

:::

::: {}
**Unsupervised Learning**:


Unsupervised learning is known for reading data without labels, e.g. pictures. 
:::

::: {}
**Reinforcement Learning**: 


Reinforcement learning is categorised as Artificial Intelligence. It is more focused on goal-directed learning from interaction than are other approaches to machine learning. As per [Sutton and Barto (2015)](https://web.stanford.edu/class/psych209/Readings/SuttonBartoIPRLBook2ndEd.pdf), the three most distinguishing features of reinforcement learning are:


    1. Being closed-loop; i.e. the learning system's actions influence its later inputs.
        
        
    2. Does not have direct instructions as to what actions to take; instead it must discover which actions yield the most reward by trying them out.
        
        
    3. Not knowing where the consequences of actions, including reward signals, play out over extended time periods.


:::

::::

**Relevant trade-offs in Machine Learning**


+ Flexibility vs. interpretability.


+ Prediction vs. inference: high predictive power does not allow for proper inference. 


+ Goodness-of-fit vs. over-fitting: how do we know *when* the fit is good?


### **An introduction to R programming**

For the purposes of this course, we will be working with the integrated development environment (IDE) Rstudio. Make sure you have [downloaded it](https://www.rstudio.com/products/rstudio/download/) and have familiarised yourself with the interface before proceeding.

**Features of R**

R is a statistical programming language. As such, it understands and categorises its input as **data types**:

(Note that R is not too strict about data types, but you need to be able to identify them to use them in math operations.)

```{r}

# Examples of data types

# Numeric (or double): 
numeric_vector <- c(1, 1.0, 65.5)
print(numeric_vector)

#Integers: 
integer_vector <- c(1L, 3L, 45L)
print(integer_vector)

#Logical (or boolean): 
boolean_vector <- c(TRUE, FALSE)
print(boolean_vector)

#Character: 
character_vector <- c("Harry Potter", "Star Wars", "Lord of the Rings")
print(character_vector)

#Null: 
NA
```

... and **data structures**. A data structure is either homogeneous (all elements are of the
same data type) or heterogeneous (elements can be of more than
one data type).

```{r}

# Examples of data structures

#Vectors: think of a row, or a column of the same element.
x <- c(1L,3L,5L,7L,9L) # we call this an integer vector
y <- c(1.3, 1, 5, 7, 11.2) # we call this a numerical (or double) vector

print(x)
print(y)

#Matrices: matrices have rows and columns containing elements of the same type.
A <- matrix(1:9, ncol=3, nrow=3, byrow= TRUE)
print(A)

#Arrays: A vector is a one-dimensional array. A matrix is a two-dimensional array.
#In short, an array is a collection of data of the same type.

n <- 5*5*3 
B <- array(1:n, c(5,5,3))
print(B)

#Lists: a list is a one-dimensional, heterogeneous data structure. Basically, 
# it is an object that stores all object types. 

my_list <- list(42, "The answer is", TRUE)
print(my_list)

#Data frames: a data frame is a list of  column vectors. Each vector must
#contain the same data type, but the different vectors can store different 
#data types. Note, however, that in a data france all vectors must have the same lenght. 

a <- 1L:5L
b <- c("a1", "b2", "c3", "d4", "e5")
c <- c(TRUE, TRUE, FALSE, TRUE, FALSE)

df <- as.data.frame(cbind(a,b,c))

str(df)
```


From here on, we will write an R script together, and learn some basic commands and tools that will allow you to explore and manipulate data. Please note that this is **not** an exhaustive tutorial. It is nonetheless a good place to start.

**1. Setting up the Rstudio working environment**


It is good practice to make sure that the working environment is empty/clean before you start running any code.
```{r}
# The r base command rm() stands for remove
rm(list = ls()) # this line indicates R to clear absolutely everything from the environment

```
Once that has been taken care of, you need to load the libraries you will be working with. While r base has a large number of commands to explore, wrangle, and manipulate data, the **open source** feature of R means that people all over the world are constantly working on packages and functions to make our lives easier. These can be used by calling the libraries in which they are stored:

```{r}
#  My personal favourite are the Tidyverse library, by Hadley Whickam, and data.table. Both are brilliant for data exploration, manipulation, and visualisation. 
library(tidyverse)
library(data.table)

```
If you're working with an imported data set, you should probably set up your working directory as well:
```{r}
setwd("~/Desktop/") #Note that
```

**2. R base commands for data set exploration**


Now that we have all the basic stuff set up, let's start with some basic r base commands that will allow us to explore our data. To do so, we will work with the toy data set *mtcars* that can be called from R without the need to upload data or call data from a website. 

```{r}

# some basics to explore your data 
str(mtcars) # show the structure of the object in a compact format
dim(mtcars) # inspect the dimension of the dataset (returns #rows, #columns)
class(mtcars) # evaluate the class of the object
typeof(mtcars) # evaluate the type of the object. typeof returns R storage type (see: https://stackoverflow.com/questions/6258004/types-and-classes-of-variables)
View(mtcars$mpg) # Try to avoid running this one in big and complex databases.
length(mtcars$mpg) # evaluate the number of elements in vector mpg
mean(mtcars$mpg) # mean of all elements in vector mpg
sum(mtcars$mpg) # sum of all elements in vector mpg (similar to a column sum)
sd(mtcars$mpg) # standard deviation
median(mtcars$mpg) # median
cor(mtcars$mpg, mtcars$wt) # default is pearson correlation, specify method within function to change it  
table(mtcars$am) #categorical data in a table: counts
prop.table(table(mtcars$am)) #categorical data in a table: proportions
```

**3. Objects and assignments**


Another important feature of the R programming language is that it is *object oriented*. That means that, for the most part, for every function used, there must be an object assigned! Let's see an example of object assignment with a bivariate linear regression model:

```{r}
ols_model <- lm(mpg ~ wt, data = mtcars) # lm stands for linear model. In parenthesis, dv first, iv after
summary(ols_model)
```
If you would like to see the results from your regression, you do not need to run it again. Instead, you print the object (ols_model) again. Similarly, you can call information stored in that object at any time, for example, the estimated coefficients:

```{r}

ols_model$coefficients
```

**4. Plotting with and without special libraries**


Now, we've loaded a couple of libraries. Why did we do that if we've only used r base commands so far? We're going to exemplify the power of libraries by drawing plots using r base, and ggplot2. Ggplot2 is the plotting function from the tidyverse, and arguably one of the best data visualisation tools across programming languages. If you'd like to read more about why that is the case, check out the [Grammar of Graphics](https://ggplot2.tidyverse.org/).

```{r}

# Plotting with R base
plot(mtcars$wt, mtcars$mpg, pch = 14, col = "grey", main ="Mileage and Weight")
abline(ols_model, col ="blue") # Note that to add a line of best fit, we had to call our previously estimate linear model, stored in the ols_model object-
# with base R, you cannot directly assign an object to a plot, you need to use...
p_rbase <- recordPlot()
# plot.new() # don't forget to clean up your device afterwards!
```

```{r}
# Plotting with ggplot2, using the grammar of graphics
# 1: linking plot to dataset, 2: definting (aes)thethic mapping, 3: use (geom)etric objects
# such as points, bars, etc. as markers, 4: the plot has layers
p_ggplot <- ggplot(data = mtcars, aes(x = wt, y = mpg, col=am )) + #the color is defined by car type (automatic 0 or manual 1)
  geom_smooth(method = "lm", col= "orange") + # no need to run a regression ex-ante to add a line of best fit 
  geom_point(alpha=0.5) + #alpha controls transparency of the geom
  theme(legend.position="none") #removing legend
print(p_ggplot)
```
Thanks to the grammar of graphics, we can continue to edit the plot after we have finished it. Perhaps we've come up with ideas to make it more stylish? Or helpful? Let's see an example:

```{r}
p_ggplot <- p_ggplot + theme(legend.position = "right") # I am saving it with the same name again, but I could easily choose another name and keep two versions of the plot. 
print(p_ggplot) # the legend we just added is NOT helpful. Why is that? 

# Remember when we talked about data types?!

class(mtcars$am) #for legends, we might prefer levels/categories 
mtcars$am <- as.factor(mtcars$am) # we have now transformed the numeric am into a factor variable 
#the importance of assigning objects :)

#  Now we can plot our scatterplot without issues
p_ggplot <- ggplot(data = mtcars, aes(x = wt, y = mpg, col = am )) +
  geom_smooth(method = "lm", col= "red") +
  geom_point(alpha=0.5) +
  theme(legend.position="right")
print(p_ggplot)

```

**5. R base capabilities and reserved names**


We've had some fun, now let's go back to some r base basics. These are going to be relevant for making algorithms of your own:

```{r}

## Arithmetic and other operations

2+2 #addition
5-2 #subtraction
33.3/2 #division
5^2 #exponentiation
200 %/% 60 #integer division  [or, how many full hours in 200 minutes?]/aka quotient
200 %% 60 #remainder [or, how many minutes are left over? ]

## Logical operators

34 < 35 #smaller than
34 < 35 | 33 #smaller than OR than (returns true if it is smaller than any one of them)
34 > 35 & 33 #bigger than AND than (returns true only if both conditions apply)
34 != 34 #negation
34 %in% 1:100 #value matching (is the object contained in the list of items?)
`%ni%` <- Negate(`%in%`) #let's create a function for NOT IN
1001 %ni% 1:100 # home-made function! :)
34==34 #evaluation

## Indexing

# The tidyverse comes with a cool Starwars dataset (or tibble, in the tidyverse language. As long as you have loaded the tidyverse library, you can use it.

starwars$name[1] #indexing: extract the first element in the name vector in the starwars dataframe
starwars[1,1] #alternative indexing: extract the element of row 1, col 1
starwars$name[2:4] # elements 2, 3, 4 of name vector
starwars[,1] #extract all elements from column 1
#notice that starwars is a tibble, not a dataframe, we'll see the implications of this later
starwars[1,] #extract all elements from row 1
starwars$height[starwars$height<150] # returns a logical vector TRUE for elements >150 in height vector
starwars[c('height', 'name')] #returns c(oncatenated) vectors

```

We are now reaching the end of this brief introduction to R and Rstudio. We will not go into the fun stuff you can do with data.table, you can find that out on your own if you need to (but know it is a powerful data wrangling package), but we will go through the *Reserved Names* in R. These are names you cannot use for your objects, because they serve a programming purpose.

```{r}
# 1. not-observed data 
# 1.1 NaN: results that cannot be reasonably defined
h <- 0/0
is.nan(h)
class(h)
print(h)

# 1.2 NA: missing data
colSums(is.na(starwars)) #How many missings in the starwars tibble*?
mean(starwars$height) # Evaluation gives NA?
is.na(starwars$height) # logical vector returns 6 true statements for is.na (coincides with colSums table)
class(starwars$height) # class evaluation returns integer...
mean(as.integer(starwars$height), na.rm = TRUE) # read as integer, ignore NAs :) (rm stands for remove)
# missing values (NA) are regarded by R as non comparables, even to themselves

# 2. if, else, ifelse, function, for, print, length (etc...)
# these words are reserved for control flow and looping 

# if else statements
meanheight <- 174.358

if ( meanheight < 175) {
  print('very tall!')
  } else {
      print('average to normal height')
    } 
# a note on using if else: else should be on the same line as if, otherwise it is not recognized

# home-made functions
f1 <- function(x){
  return(sum(x+1))
}
print(f1(5)) # returns value of x+1 when x = 5

# for loops
class(starwars$height) # if it is not integer or numeric, then transform it! 
starwars$height <- as.numeric(as.character(starwars$height))
starwars$height_judge = NA # if you're creating a new object, make sure you initialize it before running your loop

for (i in 1:length(starwars$height)) {
    
    starwars$height_judge[i] <- ifelse(starwars$height[i]<100, "Short", "Tall")
}
print(starwars[c("name", "height_judge", "height")])

```

