---
title: "Random Forests"
#author: "Dr. Stephan Dietrich & Michelle González Amador"
#date: '2022-09-21'
output: html_document
---

An introduction on Decision Trees and Random Forests was given to us by [Dr. Francisco Rosales](https://www.linkedin.com/in/fraroma/), Assistant Professor at ESAN University (Perú) and Lead Data Scientist ([BREIN](https://www.linkedin.com/company/breinhub/)). The session was not recorded, but the slides are available. 

**Lecture Slides**
```{r, echo=FALSE}
library(downloadthis)
download_link(
  link = "https://github.com/michelleg06/Machine-Learning-for-Public-Policy/blob/9f3b1dae00e0ea324c8cfd019304c78a6437afca/ml4pp_c5_fr.pdf",
  button_label = "Decision Trees Lecture Slides",
  button_type = "info",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)
```

Below you will find a simple exercise on Random Forests with an easy challenge. Please download the script and the data.

**R data**

```{r, echo=FALSE}
library(downloadthis)
download_link(
  link = "https://github.com/michelleg06/Machine-Learning-for-Public-Policy/blob/main/malawi_2005.RDS",
  button_label = "Download malawi_2005 data (RDS)",
  button_type = "info",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)

```

**R script**

```{r, echo=FALSE}
library(downloadthis)
download_link(
  link = "https://github.com/michelleg06/Machine-Learning-for-Public-Policy/blob/main/RandomForests.R",
  button_label = "Download RandomForests.R script",
  button_type = "info",
  has_icon = TRUE,
  icon = "fa fa-save",
  self_contained = FALSE
)

```
